---
title: "Chapter1"
author: "Mohammed Ali"
date: "December 26, 2017"
output: 
  pdf_document:
    toc: true
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ISLR)
library(dplyr)
library(ggplot2)
require(GGally)
require(purrr)
require(MASS)
```

# About
This document will represent the solutions for the applied part for *Intorduction to Statistical Learning* book.

# Questions
Q.8
Load college dataset and inspect it
```{r College}
data("College")
glimpse(College)
ggcorr(College[, 1:10])
ggpairs(College[, 1:10])
```

Create a new qualitative variable, called Elite to divide universities into two groups based on whether or not the proportion of students coming from the top 10% of their high school
classes exceeds 50%.

```{r}
ggplot(College, aes(y=Outstate, x=Private)) +
  geom_boxplot()
College <- College %>%
  mutate(Elite = ifelse(Top10perc > 50, "Yes", "No"))
College$Elite <- as.factor(College$Elite)
table(College$Elite)

```

```{r}
par(mfrow = c(2,2))
hist(College$Books, col = 2, xlab = "Books", ylab = "Count")
hist(College$PhD, col = 3, xlab = "PhD", ylab = "Count")
hist(College$Grad.Rate, col = 4, xlab = "Grad Rate", ylab = "Count")
hist(College$perc.alumni, col = 6, xlab = "% alumni", ylab = "Count")

data("Auto")
glimpse(Auto)
summary(Auto[-9])
map_dbl(Auto[-9], mean)
map_dbl(Auto[-9], sd)

auto_subset <-Auto[- c(10:85), -9]

map_df(auto_subset, range)
map_dbl(auto_subset, mean)
map_dbl(auto_subset, sd)

Auto$cylinders <- as.factor(Auto$cylinders)
Auto$year <- as.factor(Auto$year)
Auto$origin <- as.factor(Auto$origin)
ggcorr(Auto[-9])
ggpairs(Auto[-9])

glimpse(Boston)
ggcorr(Boston[Boston$crim < 20, ])
ggpairs(Boston[Boston$crim < 20, ])

```